\section{Strengths, Identification Risks, and Adversarial Failure}
\subsection{Why the New Formulation Is Stronger}
The new ecological-liability contract has four structural advantages over a plain hotspot-allocation approach.
\begin{itemize}
  \item It changes the decision object from a visually prominent hotspot or corridor to a priced liability surface that combines ecological consequence, pressure, uncertainty, and service drag.
  \item It treats deployment as a portfolio under ambiguity, which forces the model to care about bad scenarios and not merely about average gains.
  \item It binds the most expensive cells to an explicit minimum protection contract, which is more operationally meaningful than saying those cells are simply ``high priority.''
  \item It derives staffing from overload guarantees, so the field-force recommendation follows from a reliability argument rather than from a descriptive patrol-hours tally.
\end{itemize}

\subsection{Identification Boundary}
The framework also carries nontrivial identification risk.
\begin{itemize}
  \item If the latent incident field in Equation~\eqref{eq:prior} is built from poor access proxies, the liability surface can be spatially biased.
  \item If monitoring recency or sightline quality is mismeasured, the information-deficit premium \(U_i\) can overstate or understate ambiguity.
  \item If the scenario library \(\Omega\) is too narrow, the \(\operatorname{CVaR}\) layer will look rigorous while still missing realistic bad months.
  \item If station workload variability is underestimated, the staffing floor from Equation~\eqref{eq:rangermin} will be too low and the service covenant will be illusory.
\end{itemize}

The most important failure mode is not adversarial rerouting alone. It is \emph{adversarial substitution}: once managers protect the currently obvious ridge, pressure may migrate to a lower-intensity region whose uncertainty was previously ignored. This is exactly why the ambiguity premium is part of the objective rather than an afterthought.

\subsection{Appropriate Validation Standard}
This paper should therefore be read as a decision architecture, not as a claim of exact forecasting. The correct validation routine is to compare predicted high-liability cells with observed incident concentration, predicted contract breaches with observed patrol gaps, and predicted overload probability with realized station backlog. If those diagnostics drift, the liability field or the scenario library must be rebuilt before the deployment recommendations are trusted again.
