\section{Decision Answer Under the New Contract}
This paper reaches a different strategic answer by changing both the narrative and the mathematics. Etosha should not be managed as a coverage exercise, and it should not be reduced to a single corridor-sealing controller. It should be managed as an ecological-liability contract under ambiguity.

That reformulation changes the architecture, the algorithms, and the conclusions at the same time. The architecture is now a liability-field estimator, an ambiguity-aware portfolio optimizer, and a workload-guarantee layer. The algorithms first identify contract cells, then optimize mean loss plus \(\operatorname{CVaR}\), and finally size crews so that station overload remains acceptable. The conclusion is no longer ``cover the corridor as much as possible.'' It is ``guarantee minimum protection on the most expensive cells, use the remaining budget to suppress tail risk, and staff stations to a workload covenant rather than to an average-hour target.''

Direct answers to Tasks 1--6 are therefore:
\begin{itemize}
  \item Task 1: define protection as the reduction of unresolved ecological liability under an explicit service covenant, and diagnose a narrow liability ridge as the dominant management problem.
  \item Task 2: deploy a contract-based portfolio of patrol templates, sentinel devices, and UAV reserve rather than a broad uniform patrol layer.
  \item Task 3: maintain roughly 34 field rangers under baseline conditions to keep overload probability within the stated covenant.
  \item Task 4: recognize team depth, contract-cell uptime, and revisit discipline as the dominant structural levers.
  \item Task 5: treat liability-field misspecification, scenario misspecification, and underestimated workload volatility as the principal technical risks.
  \item Task 6: transfer the framework by rebuilding liability morphology and station workload structure for the new park.
\end{itemize}

The broader lesson is methodological. Conservation planning improves when it stops asking where effort can be seen and starts asking where unresolved ecological liability remains too expensive to tolerate. For Etosha, that means defending a contract, not decorating a map.

\newpage
