\begin{thebibliography}{99}

\bibitem[Etosha National Park(2026)]{etoshaOfficial}
Etosha National Park.
\newblock \emph{Etosha National Park}.
\newblock 2026.
\newblock \url{https://www.etoshanationalpark.org/}. Accessed March 4, 2026.

\bibitem[Etosha National Park(2025)]{etoshaMap2025}
Etosha National Park.
\newblock \emph{Etosha National Park Map}.
\newblock 2025.
\newblock \url{https://etoshanationalpark.co.za/wp-content/uploads/sites/16/2025/03/Etosha-National-Park-Map.jpg}. Accessed March 5, 2026.

\bibitem[van Wijngaarden(1985)]{etoshaEcology}
Fransje van Wijngaarden.
\newblock \emph{Environmental and Vegetation Gradients in the Etosha National Park}.
\newblock 1985.
\newblock \url{https://www.scielo.org.za/scielo.php?script=sci_arttext&pid=S0075-64581985000100009}. Accessed March 4, 2026.

\bibitem[Fang et~al.(2016)Fang, Stone, and Tambe]{fang2016paws}
Fei Fang, Peter Stone, and Milind Tambe.
\newblock When security games go green: Designing defender strategies to prevent poaching and illegal fishing.
\newblock In \emph{Proceedings of the Twenty-Fifth International Joint Conference on Artificial Intelligence}, pages 2589--2595, 2016.

\bibitem[U.S. Geological Survey(2018)]{usgsDetection}
U.S. Geological Survey.
\newblock \emph{Some Anti-Poaching Patrols Are Much More Efficient Than Others}.
\newblock 2018.
\newblock \url{https://www.usgs.gov/news/featured-story/some-anti-poaching-patrols-are-much-more-efficient-others}. Accessed March 4, 2026.

\bibitem[National Park Service(2026)]{yellowstoneOfficial}
National Park Service.
\newblock \emph{Yellowstone National Park}.
\newblock 2026.
\newblock \url{https://www.nps.gov/yell/index.htm}. Accessed March 4, 2026.

\bibitem[Parks Australia(2026)]{kakaduOfficial}
Parks Australia.
\newblock \emph{Kakadu National Park}.
\newblock 2026.
\newblock \url{https://parksaustralia.gov.au/kakadu/}. Accessed March 4, 2026.

\bibitem[Dantzig(1963)]{dantzig1963}
George B. Dantzig.
\newblock \emph{Linear Programming and Extensions}.
\newblock Princeton University Press, Princeton, NJ, 1963.

\bibitem[Birge and Louveaux(2011)]{birge2011}
John R. Birge and Francois Louveaux.
\newblock \emph{Introduction to Stochastic Programming}.
\newblock 2nd edition, Springer, New York, 2011.

\bibitem[Law(2015)]{law2015}
Averill M. Law.
\newblock \emph{Simulation Modeling and Analysis}.
\newblock 5th edition, McGraw-Hill Education, New York, 2015.

\bibitem[Banks et~al.(2010)Banks, Carson, Nelson, and Nicol]{banks2010}
Jerry Banks, John S. Carson, Barry L. Nelson, and David M. Nicol.
\newblock \emph{Discrete-Event System Simulation}.
\newblock 5th edition, Prentice Hall, Upper Saddle River, NJ, 2010.

\bibitem[Bonabeau(2002)]{bonabeau2002}
Eric Bonabeau.
\newblock Agent-based modeling: Methods and techniques for simulating human systems.
\newblock \emph{Proceedings of the National Academy of Sciences}, 99(suppl. 3):7280--7287, 2002.

\bibitem[Macal and North(2010)]{macal2010}
Charles M. Macal and Michael J. North.
\newblock Tutorial on agent-based modelling and simulation.
\newblock \emph{Journal of Simulation}, 4(3):151--162, 2010.

\bibitem[Cressie(1993)]{cressie1993}
Noel A. C. Cressie.
\newblock \emph{Statistics for Spatial Data}.
\newblock Revised edition, Wiley, New York, 1993.

\bibitem[Anselin(1995)]{anselin1995}
Luc Anselin.
\newblock Local indicators of spatial association---LISA.
\newblock \emph{Geographical Analysis}, 27(2):93--115, 1995.

\bibitem[Diggle and Ribeiro(2013)]{diggle2013}
Peter J. Diggle and Paulo J. Ribeiro.
\newblock \emph{Model-based Geostatistics}.
\newblock 2nd edition, Springer, New York, 2013.

\bibitem[Weiss et~al.(2018)]{weiss2018}
D. J. Weiss, A. Nelson, H. S. Gibson, W. Temperley, S. Peedell, A. Lieber, M. Hancher, E. Poyart, S. Belchior, N. Fullman, B. Mappin, U. Dalrymple, J. Rozier, T. C. D. Lucas, R. E. Howes, L. S. Tusting, S. Y. Kang, E. Cameron, D. Bisanzio, K. E. Battle, S. Bhatt, and P. W. Gething.
\newblock A global map of travel time to cities to assess inequalities in accessibility in 2015.
\newblock \emph{Nature}, 553(7688):333--336, 2018.

\bibitem[Watson et~al.(2014)Watson, Dudley, Segan, and Hockings]{watson2014}
James E. M. Watson, Nigel Dudley, Daniel B. Segan, and Marc Hockings.
\newblock The performance and potential of protected areas.
\newblock \emph{Nature}, 515(7525):67--73, 2014.

\bibitem[Geldmann et~al.(2013)Geldmann, Barnes, Coad, Craigie, Hockings, and Burgess]{geldmann2013}
Jonas Geldmann, Megan Barnes, Lauren Coad, Ian D. Craigie, Marc Hockings, and Neil D. Burgess.
\newblock Effectiveness of terrestrial protected areas in reducing habitat loss and population declines.
\newblock \emph{Biological Conservation}, 161:230--238, 2013.

\bibitem[Jones et~al.(2018)Jones, Venter, Fuller, Allan, Maxwell, Negret, and Watson]{jones2018}
Kendall R. Jones, Oscar Venter, Richard A. Fuller, James R. Allan, Sean L. Maxwell, Pablo J. Negret, and James E. M. Watson.
\newblock One-third of global protected land is under intense human pressure.
\newblock \emph{Science}, 360(6390):788--791, 2018.

\bibitem[Coad et~al.(2019)Coad, Watson, Geldmann, Burgess, Leverington, Hockings, Knights, and Di Marco]{coad2019}
Lauren Coad, James E. M. Watson, Jonas Geldmann, Neil D. Burgess, Fiona Leverington, Marc Hockings, Katie Knights, and Moreno Di Marco.
\newblock Widespread shortfalls in protected area resourcing undermine efforts to conserve biodiversity.
\newblock \emph{Frontiers in Ecology and the Environment}, 17(5):259--264, 2019.

\bibitem[Moore et~al.(2018)Moore, Mulindahabi, Masozera, Nichols, Hines, Turikunkiko, and Oli]{moore2018}
Jennifer F. Moore, Felix Mulindahabi, Michel K. Masozera, James D. Nichols, James E. Hines, Ezechiel Turikunkiko, and Madan K. Oli.
\newblock Are ranger patrols effective in reducing poaching-related threats within protected areas?
\newblock \emph{Journal of Applied Ecology}, 55(1):99--107, 2018.

\bibitem[IUCN(2026)]{iucnRedList2026}
IUCN.
\newblock \emph{The IUCN Red List of Threatened Species}.
\newblock 2026.
\newblock \url{https://www.iucnredlist.org/}. Accessed March 6, 2026.

\bibitem[UNODC(2024)]{unodc2024wildlife}
United Nations Office on Drugs and Crime.
\newblock \emph{World Wildlife Crime Report 2024: Trafficking in Protected Species}.
\newblock 2024.
\newblock \url{https://www.unodc.org/unodc/en/data-and-analysis/wildlife.html}. Accessed March 6, 2026.

\end{thebibliography}
